---
title: "Network Analysis Homework: Hollywood Actors Graph"
author: "Aiden Dickson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## I. Getting Ready

```{r}
set.seed(1111)
library(igraph)
library(ggraph)
library(corrplot)
```

------------------------------------------------------------------------

## Task 1: Re-read the Graph from GitHub (Cloud)

```{r}
fileLink = 'https://github.com/AidenFlowers/assign2-computation-socsci-methods-dacss-690c/raw/main/homework-resources/hollywood.graphml'
actors = read_graph(fileLink, format = 'graphml')
head(actors)
```

------------------------------------------------------------------------

## Task 2: Compute Node Centrality Measures

```{r}
eigen = eigen_centrality(actors)$vector
close = closeness(actors, normalized = T)
betw = betweenness(actors, normalized = T)

DFCentrality = as.data.frame(cbind(eigen, close, betw), stringsAsFactors = F)
names(DFCentrality) = c('Eigenvector', 'Closeness', 'Betweenness')
DFCentrality$person = row.names(DFCentrality)

DFCentrality
```

------------------------------------------------------------------------

## Task 3: Correlation Plot

```{r}
cor_matrix = cor(DFCentrality[, c("Eigenvector", "Closeness", "Betweenness")])

corrplot(cor_matrix, 
         method = "circle", 
         type = "upper", 
         addCoef.col = "black", 
         tl.col = "black",
         title = "Correlation of Centrality Measures",
         mar = c(0, 0, 2, 0))
```

### Interpretation

-   **Eigenvector vs Closeness (-0.68):** Strong negative correlation. Actors connected to other highly-connected actors form tight clusters but are distant from other groups.
-   **Eigenvector vs Betweenness (-0.25):** Weak negative correlation. Being well-connected to important actors doesn't strongly predict being a bridge between groups.
-   **Closeness vs Betweenness (+0.77):** Strong positive correlation. Actors who can reach others quickly also tend to lie on shortest paths between others.

This pattern suggests a network with distinct clusters where some actors dominate one cluster while others serve as bridges.

------------------------------------------------------------------------

## Task 4: Is it Possible Communities are Present?

```{r}
 

RandomNet = sample_gnp(length(V(actors)), p = edge_density(actors))
ratio = transitivity(actors) / transitivity(RandomNet)
cat("Actors transitivity:", transitivity(actors), "\n")
cat("Random transitivity:", transitivity(RandomNet), "\n")
cat("Resultant Ratio:", sprintf("%.3f", ratio), "\n")
```

### Interpretation

-   **Resultant Ratio** The resultant ratio is 0.875, which is less than 1. This means the Hollywood actors network has slightly less clustering than a comparable random network. Based on the transitivity test alone, there is no strong evidence for community structure. However, this test has limitations for small networks (11 nodes). \*As we will see, the modularity scores from community detection algorithms (Girvan-Newman and Louvain, both \> 0.4) provide stronger evidence that meaningful communities do exist.

------------------------------------------------------------------------

## Task 5: Community Detection

### Girvan-Newman Algorithm

```{r}
partition_girvanNewman_best = cluster_edge_betweenness(actors)
gn_modularity = max(partition_girvanNewman_best$modularity)
cat("Girvan-Newman Modularity:", gn_modularity, "\n")
```

### Louvain Algorithm

```{r}
partition_louvain_optimal = cluster_louvain(actors)
louvain_modularity = max(partition_louvain_optimal$modularity)
cat("Louvain Modularity:", louvain_modularity, "\n")
```

### Plot the Best Result

```{r}
V(actors)$louvain = as.factor(membership(partition_louvain_optimal))
V(actors)$girvanNewman = as.factor(membership(partition_girvanNewman_best))

best_algorithm = ifelse(gn_modularity > louvain_modularity, "Girvan-Newman", "Louvain")
best_partition = ifelse(gn_modularity > louvain_modularity, "girvanNewman", "louvain")
best_modularity = max(gn_modularity, louvain_modularity)

base = ggraph(graph = actors, layout = 'fr') + geom_edge_link(alpha = 0.2)

base +
  geom_node_point(aes(color = get(best_partition)), show.legend = TRUE, size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Community Detection:", best_algorithm, "(Best Result)"),
       subtitle = paste("Modularity:", round(best_modularity, 3)),
       color = "Community") +
  theme_graph()
```
